---
title: "BighornWeatherEnvironmentAnalysis"
author: "DanielleGlass"
date: "3/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I decided to use Daymet remote sensing data for the weather data to compared against bighorn water use. Since the Daymet data is calculated for each square kilometer on the Earth's surface, I had to decide which square kilometer to obtain Daymet data for in each mountain range. I choose locations that were at approximately the mean elevation for the range, and that were somewhat in the middle of known water sources so as to closely model what bighorn visiting were experiencing weather-wise. It is worth noting that the elevations cho The elevations choosen are below:

Using CalTopo: 
Marbles: 597' - 3840' (181 m - 1171 m) --> 676
South Bristols: 703' - 3269' (214 m - 997 m) --> 605.5
NROM: 1800'-6095' (548 m - 1858 m) --> 1203
Castle Piute: 4000'-5185' (1219 m - 1580 m) --> 1399.5
Nopah: 1392'-5249' (424 m - 1599 m) --> 1011.5

Marbles: 34.6298, -115.5856


```{r}
#Marbles
marbdaymet <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/Marbdaymet11193_lat_34.6298_lon_-115.5856.csv", sep = ",", header = TRUE)
# 7/25/19 only day it rained in the Marbles all summer
# Cut to the days when cameras were running
marbdaymet$Date <- as.character(marbdaymet$Date)
marbdaymet$Date <- as.Date(marbdaymet$Date)
marbdaymetcut <- marbdaymet[(marbdaymet$Date >= as.Date("2019-05-18") & marbdaymet$Date <= as.Date("2019-05-24")) | (marbdaymet$Date >= as.Date("2019-06-05") & marbdaymet$Date <= as.Date("2019-06-11")) | (marbdaymet$Date >= as.Date("2019-06-19") & marbdaymet$Date <= as.Date("2019-06-22")) | (marbdaymet$Date >= as.Date("2019-06-26") & marbdaymet$Date <= as.Date("2019-06-29")) | (marbdaymet$Date >= as.Date("2019-07-11") & marbdaymet$Date <= as.Date("2019-07-13")) |  (marbdaymet$Date >= as.Date("2019-07-18") & marbdaymet$Date <= as.Date("2019-07-24")) | (marbdaymet$Date >= as.Date("2019-08-01") & marbdaymet$Date <= as.Date("2019-08-03")) | (marbdaymet$Date >= as.Date("2019-08-07") & marbdaymet$Date <= as.Date("2019-08-09")), ]
# Read in NAs for days when the cameras weren't running to be able to compare against/plot
marbwthrna <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/MarbNAweatherenvironmentalvariables.csv", sep = ",", header = TRUE)
marbwthrna$Date <- as.Date(marbwthrna$Date)
# Plotting max temperature
marbdatetempmax <- data.frame(marbdaymetcut$Date, marbdaymetcut$tmax..deg.c.)
names(marbdatetempmax)[names(marbdatetempmax) == "marbdaymetcut.Date"] <- "Date"
names(marbdatetempmax)[names(marbdatetempmax) == "marbdaymetcut.tmax..deg.c."] <- "Tmax"
names(marbwthrna)[names(marbwthrna) == "Nas"] <- "Tmax"
marbdatetempmaxna <- rbind(marbdatetempmax, marbwthrna)
plot(marbdatetempmaxna$Tmax ~ marbdatetempmaxna$Date, type = "b", xlab = "Date", ylab = "Maximum Daily Temperature (°C)", main = "Marbles Maximum Daily Temperature")
# It is worth noting that on this plot, there are lines connecting over NA dates.
# Plotting min temperature



```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
