---
title: "BighornWeatherEnvironmentAnalysis"
author: "DanielleGlass"
date: "3/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I decided to use Daymet remote sensing data for the weather data to compare against bighorn water use. Since the Daymet data is calculated for each square kilometer on the Earth's surface, I had to decide which square kilometer to obtain Daymet data for in each mountain range. I choose locations that were at approximately the mean elevation for the range, and that were somewhat in the middle of known water sources so as to closely model what bighorn visiting were experiencing weather-wise. It is worth noting that the elevations choosen are below:

Using CalTopo: 
Marbles: 597' - 3840' (181 m - 1171 m) --> 676
South Bristols: 703' - 3269' (214 m - 997 m) --> 605.5
NROM: 1800'-6095' (548 m - 1858 m) --> 1203
Castle Piute: 4000'-5185' (1219 m - 1580 m) --> 1399.5
Nopah: 1392'-5249' (424 m - 1599 m) --> 1011.5

Marbles: 34.6298, -115.5856
South Bristols: 34.5937, -115.6823
CMPR: 35.2795, -115.0990
Nopah: 36.0105, -116.1015

First, I uploaded the Daymet NDVI data. Then I cut the visitation data down to mountain range, date, percent of sheep who visited that day, and total number of visits/total number of sheep. Then I merged these two datasets based on mountain range and date.

```{r}
allrangeenvvar <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/AllRangesDaymetNA.csv", sep = ",", header = TRUE)

# Percent of Individual Sheep Who visited on a given day, in a given range
# allindsheepvisitperday
# marbindsheepvisitperday1, sbrsindsheepvisitperday2, nopaindsheepvisitperday2, cmprindsheepvisitperday2

# Mean number of Visits/Day/Collared Bighorn
# allindrd11
# marbindrd11, sbrsrd11, nopard11, cmprrd11

meanpercentbighorn <- merge(x = allindsheepvisitperday, y = allindrd11, all.x = TRUE)
meanpercentbighorn1 <- meanpercentbighorn[with(meanpercentbighorn, order(MtnRange, Date)), ]
# There were some duplicated rows for the Marbles but I could not for the life of me figure out why. So I deleted them.
rownames(meanpercentbighorn1) <- seq(length=nrow(meanpercentbighorn1))
meanpercentbighorn2 <- meanpercentbighorn1[-c(3, 5, 7, 8, 11, 12, 26, 28, 29, 31, 33, 36, 37, 46, 48, 51, 52, 58, 59, 62, 64), ]
names(allrangeenvvar)[names(allrangeenvvar) == "MountainRange"] <- "MtnRange"
allenvvisit <- merge(x = meanpercentbighorn2, y = allrangeenvvar, all = TRUE)
allenvvisit1 <- allenvvisit[with(allenvvisit, order(MtnRange, Date)), ]
```
First the distributions of the dependent variables.

I checked to make sure that variables were more of less normally distributed.





```{r}
# Checking distribution of the dependent variables
qqnorm(allenvvisit1$Percentvisited)
qqline(allenvvisit1$Percentvisited)
hist(allenvvisit1$Percentvisited)


require(car)
require(MASS)
allenvvisit1$Percentvisiteddist <- allenvvisit1$Percentvisited + 1
qqp(allenvvisit1$Percentvisiteddist, "norm")
qqp(allenvvisit1$Percentvisiteddist, "lnorm")
# nbinom <- fitdistr(allenvvisit1$Percentvisiteddist, "Negative Binomial")
# qqp(allenvvisit1$Percentvisited, "nbinom", size = nbinom$estimate[[1]], mu = nbinom$estimate[[2]])
# poissonpv <- fitdistr(allenvvisit1$Percentvisited, "Poisson")
# qqp(allenvvisit1$Percentvisited, poissonpv$estimate)

# Looks linear enough to me, though the first quartile is wonky because of a lot of zeros. Is this good enough to be considered normal?
qqnorm(allenvvisit1$Meannumvisits)
qqline(allenvvisit1$Meannumvisits)
hist(allenvvisit1$Meannumvisits)
# Same thing on this one. IS THIS NORMAL ENOUGH? NO.

# Checked for normal distribution
qqnorm(allenvvisit1$dayl..s.)
qqline(allenvvisit1$dayl..s.)
hist(allenvvisit1$dayl..s.)
# Day length is normally distributed.
qqnorm(allenvvisit1$prcp..mm.day.)
qqline(allenvvisit1$prcp..mm.day.)
hist(allenvvisit1$prcp..mm.day.)
# Normality is not super relevant when basically 100% of this is zero.
qqnorm(allenvvisit1$srad..W.m.2.)
qqline(allenvvisit1$srad..W.m.2.)
hist(allenvvisit1$srad..W.m.2.)
# Solar radiation is normally distributed.
qqnorm(allenvvisit1$tmax..deg.c.)
qqline(allenvvisit1$tmax..deg.c.)
hist(allenvvisit1$tmax..deg.c.)
# Max temp is normally distributed.
qqnorm(allenvvisit1$tmin..deg.c.)
qqline(allenvvisit1$tmin..deg.c.)
hist(allenvvisit1$tmin..deg.c.)
# Min temp is more or less normally distributed.
qqnorm(allenvvisit1$vp..Pa.)
qqline(allenvvisit1$vp..Pa.)
hist(allenvvisit1$vp..Pa.)
# Vapor pressure is normally distributed
qqnorm(allenvvisit1$MeanNDVI)
qqline(allenvvisit1$MeanNDVI)
hist(allenvvisit1$MeanNDVI)
# NDVI is normally distributed

# percentvisitedglmall <- glm(Percentvisited ~ (MtnRange + dayl..s. + prcp..mm.day. + srad..W.m.2. + tmax..deg.c. + tmin..deg.c. + vp..Pa. + MeanNDVI), data = allenvvisit1, family = poisson)

```
Plotting the independent variables through time




```{r}
marballen <- allenvvisit1[allenvvisit1$MtnRange == "Marb", ]
sbrsallen <- allenvvisit1[allenvvisit1$MtnRange == "BrsS", ]
nopaallen <- allenvvisit1[allenvvisit1$MtnRange == "Nopa", ]
cmprallen <- allenvvisit1[allenvvisit1$MtnRange == "CMPR", ]

# Max temperature
library(ggplot2)
tmaxplot <- ggplot() +
  geom_line(data = marballen, aes(x = Date, y = tmax..deg.c., color = "red"))+
  geom_point(data = marballen, aes(x = Date, y = tmax..deg.c., color = "red"))+
  geom_line(data = sbrsallen, aes(x = Date, y = tmax..deg.c., color = "blue"))+
  geom_point(data = sbrsallen, aes(x = Date, y = tmax..deg.c., color = "blue"))+
  geom_line(data = nopaallen, aes(x = Date, y = tmax..deg.c., color = "green"))+
  geom_point(data = nopaallen, aes(x = Date, y = tmax..deg.c., color = "green"))+
  geom_line(data = cmprallen, aes(x = Date, y = tmax..deg.c., color = "orange"))+
  geom_point(data = cmprallen, aes(x = Date, y = tmax..deg.c., color = "orange"))+
  ylab('Maximum Daily Temperature (°C)')+
  scale_color_discrete(labels = c("Marbles", "South Bristols", "Nopah", "Castle Piute"))+
  ggtitle("Maximum Daily Temperature")
print(tmaxplot)

# Minimum Temperature

tminplot <- ggplot() +
  geom_line(data = marballen, aes(x = Date, y = tmin..deg.c., color = "red"))+
  geom_point(data = marballen, aes(x = Date, y = tmin..deg.c., color = "red"))+
  geom_line(data = sbrsallen, aes(x = Date, y = tmin..deg.c., color = "blue"))+
  geom_point(data = sbrsallen, aes(x = Date, y = tmin..deg.c., color = "blue"))+
  geom_line(data = nopaallen, aes(x = Date, y = tmin..deg.c., color = "green"))+
  geom_point(data = nopaallen, aes(x = Date, y = tmin..deg.c., color = "green"))+
  geom_line(data = cmprallen, aes(x = Date, y = tmin..deg.c., color = "orange"))+
  geom_point(data = cmprallen, aes(x = Date, y = tmin..deg.c., color = "orange"))+
  ylab('Minimum Daily Temperature (°C)')+
  scale_color_discrete(labels = c("Marbles", "South Bristols", "Nopah", "Castle Piute"))+
  ggtitle("Minimum Daily Temperature")
print(tminplot)

# Mean NDVI

ndviplot <- ggplot() +
  geom_line(data = marballen, aes(x = Date, y = MeanNDVI, color = "red"))+
  geom_point(data = marballen, aes(x = Date, y = MeanNDVI, color = "red"))+
  geom_line(data = sbrsallen, aes(x = Date, y = MeanNDVI, color = "blue"))+
  geom_point(data = sbrsallen, aes(x = Date, y = MeanNDVI, color = "blue"))+
  geom_line(data = nopaallen, aes(x = Date, y = MeanNDVI, color = "green"))+
  geom_point(data = nopaallen, aes(x = Date, y = MeanNDVI, color = "green"))+
  geom_line(data = cmprallen, aes(x = Date, y = MeanNDVI, color = "orange"))+
  geom_point(data = cmprallen, aes(x = Date, y = MeanNDVI, color = "orange"))+
  ylab('Mean NDVI')+
  scale_color_discrete(labels = c("Marbles", "South Bristols", "Nopah", "Castle Piute"))+
  ggtitle("Mean NDVI")
print(ndviplot)

# Water Vapor Pressure
vpplot <- ggplot() +
  geom_line(data = marballen, aes(x = Date, y = vp..Pa., color = "red"))+
  geom_point(data = marballen, aes(x = Date, y = vp..Pa., color = "red"))+
  geom_line(data = sbrsallen, aes(x = Date, y = vp..Pa., color = "blue"))+
  geom_point(data = sbrsallen, aes(x = Date, y = vp..Pa., color = "blue"))+
  geom_line(data = nopaallen, aes(x = Date, y = vp..Pa., color = "green"))+
  geom_point(data = nopaallen, aes(x = Date, y = vp..Pa., color = "green"))+
  geom_line(data = cmprallen, aes(x = Date, y = vp..Pa., color = "orange"))+
  geom_point(data = cmprallen, aes(x = Date, y = vp..Pa., color = "orange"))+
  ylab('Daily Average Partial Pressure of Water Vapor (Pascals)')+
  scale_color_discrete(labels = c("Marbles", "South Bristols", "Nopah", "Castle Piute"))+
  ggtitle("Water Vapor Pressure")
print(vpplot)

# Day Length
daylength <- ggplot() +
  geom_line(data = marballen, aes(x = Date, y = dayl..s., color = "red"))+
  geom_point(data = marballen, aes(x = Date, y = dayl..s., color = "red"))+
  geom_line(data = sbrsallen, aes(x = Date, y = dayl..s., color = "blue"))+
  geom_point(data = sbrsallen, aes(x = Date, y = dayl..s., color = "blue"))+
  geom_line(data = nopaallen, aes(x = Date, y = dayl..s., color = "green"))+
  geom_point(data = nopaallen, aes(x = Date, y = dayl..s., color = "green"))+
  geom_line(data = cmprallen, aes(x = Date, y = dayl..s., color = "orange"))+
  geom_point(data = cmprallen, aes(x = Date, y = dayl..s., color = "orange"))+
  ylab('Day Length (seconds/day)')+
  scale_color_discrete(labels = c("Marbles", "South Bristols", "Nopah", "Castle Piute"))+
  ggtitle("Day Length")
print(daylength)

# Precipitation (mm/day)
dayprecip <- ggplot() +
  geom_line(data = marballen, aes(x = Date, y = prcp..mm.day., color = "red"))+
  geom_point(data = marballen, aes(x = Date, y = prcp..mm.day., color = "red"))+
  geom_line(data = sbrsallen, aes(x = Date, y = prcp..mm.day., color = "blue"))+
  geom_point(data = sbrsallen, aes(x = Date, y = prcp..mm.day., color = "blue"))+
  geom_line(data = nopaallen, aes(x = Date, y = prcp..mm.day., color = "green"))+
  geom_point(data = nopaallen, aes(x = Date, y = prcp..mm.day., color = "green"))+
  geom_line(data = cmprallen, aes(x = Date, y = prcp..mm.day., color = "orange"))+
  geom_point(data = cmprallen, aes(x = Date, y = prcp..mm.day., color = "orange"))+
  ylab('Precipitation (mm/day)')+
  scale_color_discrete(labels = c("Marbles", "South Bristols", "Nopah", "Castle Piute"))+
  ggtitle("Precipitation")
print(dayprecip)

# Shortwave Radiation
swradiation <- ggplot() +
  geom_line(data = marballen, aes(x = Date, y = srad..W.m.2., color = "red"))+
  geom_point(data = marballen, aes(x = Date, y = srad..W.m.2., color = "red"))+
  geom_line(data = sbrsallen, aes(x = Date, y = srad..W.m.2., color = "blue"))+
  geom_point(data = sbrsallen, aes(x = Date, y = srad..W.m.2., color = "blue"))+
  geom_line(data = nopaallen, aes(x = Date, y = srad..W.m.2., color = "green"))+
  geom_point(data = nopaallen, aes(x = Date, y = srad..W.m.2., color = "green"))+
  geom_line(data = cmprallen, aes(x = Date, y = srad..W.m.2., color = "orange"))+
  geom_point(data = cmprallen, aes(x = Date, y = srad..W.m.2., color = "orange"))+
  ylab('Shortwave Radiation (watts/m2)')+
  scale_color_discrete(labels = c("Marbles", "South Bristols", "Nopah", "Castle Piute"))+
  ggtitle("Incident Shortwave Radiation Flux Density")
print(swradiation)

#Percentvisited - already made other figure of this
percentvisitedplotty <- ggplot() +
  geom_line(data = marballen, aes(x = Date, y = Percentvisited, color = "red"))+
  geom_point(data = marballen, aes(x = Date, y = Percentvisited, color = "red"))+
  geom_line(data = sbrsallen, aes(x = Date, y = Percentvisited, color = "blue"))+
  geom_point(data = sbrsallen, aes(x = Date, y = Percentvisited, color = "blue"))+
  geom_line(data = nopaallen, aes(x = Date, y = Percentvisited, color = "green"))+
  geom_point(data = nopaallen, aes(x = Date, y = Percentvisited, color = "green"))+
  geom_line(data = cmprallen, aes(x = Date, y = Percentvisited, color = "orange"))+
  geom_point(data = cmprallen, aes(x = Date, y = Percentvisited, color = "orange"))+
  ylab('Percent of Collared Bighorn Visiting Water')+
  scale_color_discrete(labels = c("Marbles", "South Bristols", "Nopah", "Castle Piute"))+
  ggtitle("Percent of Collared Bighorn Visiting Water by Day")
print(percentvisitedplotty)



```



```{r}
# logpoparea <- glm(Population ~ log(Area), data = state, family )
# summary(logpoparea)
```

I then performed 2 mixed-effect general linear models to determine what environmental variables most explain the percent of individual sheep visiting water (on a given day, in a given mountain range) and the mean number of visits/day/collared bighorn.

My independent variables were: mountain range, duration of the daylight period (dayl, in seconds per day), total daily precipitation (prcp, in mm/day), shortwave radiation(srad, Incident shortwave radiation flux density in watts per square meter, taken as an average over the daylight period of the day), maximum air temperature (tmax, daily maximum 2-meter air temperature in degrees Celsius), minimum air temperature (tmin, daily minimum 2-meter air temperature in degrees Celsius), water vapor pressure (vp, daily average partial pressure of water vapor in pascals), and normalized difference vegetation index (average of the NDVI index of 250 m raster cells in each mountain range).


```{r}
# Number of Visits to Water
library(GLMMadaptive)
allenvvisit1$NumIndVisits <- (allenvvisit1$Meannumvisits * allenvvisit1$TotalNumSheep)
allenvvisit1$Date <- as.factor(allenvvisit1$Date)

#Number of Individual Sheep that Visited Water



# 
fm1 <- mixed_model(NumIndSheep ~ dayl..s. + prcp..mm.day. + srad..W.m.2. + tmax..deg.c. + tmin..deg.c. + vp..Pa. + MeanNDVI, random = ~ 1 | (MtnRange + Date + TotalNumSheep),  data = allenvvisit1, family = zi.poisson(), zi_fixed = ~ dayl..s. + prcp..mm.day. + srad..W.m.2. + tmax..deg.c. + tmin..deg.c. + vp..Pa. + MeanNDVI)


# Linear Mixed Model
library(lme4)
library(car)
Percentvisitedlmer <- lmer(Percentvisited ~ dayl..s. + prcp..mm.day. + srad..W.m.2. + tmax..deg.c. + tmin..deg.c. + vp..Pa. + MeanNDVI + (1|MtnRange) + (1|Date), data = allenvvisit1, REML = FALSE)
# Warning - predictor variables are on very different scales - consider rescaling?
summary(Percentvisitedlmer)
# Variance of fixed effect = 0.01543 --> pretty close to zero so can use normal glm?
plot(Percentvisitedlmer)
# Plot indicates that there is something going on that my model is not accounting for?
qqnorm(resid(Percentvisitedlmer))
qqline(resid(Percentvisitedlmer))
# qqnorm looks good
Anova(Percentvisitedlmer)

glm(Percentvisited ~ tmax..deg.c., data = allenvvisit1)
```
BEFORE INPUTTING COMPLETED EXCEL SPREADSHEET:

```{r}
#Marbles
marbdaymet <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/Marbdaymet11193_lat_34.6298_lon_-115.5856.csv", sep = ",", header = TRUE)
# 7/25/19 only day it rained in the Marbles all summer
# Cut to the days when cameras were running
marbdaymet$Date <- as.character(marbdaymet$Date)
marbdaymet$Date <- as.Date(marbdaymet$Date)
marbdaymetcut <- marbdaymet[(marbdaymet$Date >= as.Date("2019-05-18") & marbdaymet$Date <= as.Date("2019-05-24")) | (marbdaymet$Date >= as.Date("2019-06-05") & marbdaymet$Date <= as.Date("2019-06-11")) | (marbdaymet$Date >= as.Date("2019-06-19") & marbdaymet$Date <= as.Date("2019-06-22")) | (marbdaymet$Date >= as.Date("2019-06-26") & marbdaymet$Date <= as.Date("2019-06-29")) | (marbdaymet$Date >= as.Date("2019-07-11") & marbdaymet$Date <= as.Date("2019-07-13")) |  (marbdaymet$Date >= as.Date("2019-07-18") & marbdaymet$Date <= as.Date("2019-07-24")) | (marbdaymet$Date >= as.Date("2019-08-01") & marbdaymet$Date <= as.Date("2019-08-03")) | (marbdaymet$Date >= as.Date("2019-08-07") & marbdaymet$Date <= as.Date("2019-08-09")), ]
# Read in NAs for days when the cameras weren't running to be able to compare against/plot
marbwthrna <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/MarbNAweatherenvironmentalvariables.csv", sep = ",", header = TRUE)
marbwthrna$Date <- as.Date(marbwthrna$Date)
# Plotting max temperature
marbdatetempmax <- data.frame(marbdaymetcut$MountainRange, marbdaymetcut$Date, marbdaymetcut$tmax..deg.c.)
names(marbdatetempmax)[names(marbdatetempmax) == "marbdaymetcut.MountainRange"] <- "MountainRange"
names(marbdatetempmax)[names(marbdatetempmax) == "marbdaymetcut.Date"] <- "Date"
names(marbdatetempmax)[names(marbdatetempmax) == "marbdaymetcut.tmax..deg.c."] <- "Tmax"
names(marbwthrna)[names(marbwthrna) == "Nas"] <- "Tmax"
marbdatetempmaxna <- rbind(marbdatetempmax, marbwthrna)
# plot(marbdatetempmaxna$Tmax ~ marbdatetempmaxna$Date, type = "b", xlab = "Date", ylab = "Maximum Daily Temperature (°C)", main = "Marbles Maximum Daily Temperature")
# It is worth noting that on this plot, there are lines connecting over NA dates.
# Plotting min temperature
marbdatetempmin <- data.frame(marbdaymetcut$Date, marbdaymetcut$tmin..deg.c.)
names(marbdatetempmin)[names(marbdatetempmin) == "marbdaymetcut.Date"] <- "Date"
names(marbdatetempmin)[names(marbdatetempmin) == "marbdaymetcut.tmin..deg.c."] <- "Tmin"
names(marbwthrna)[names(marbwthrna) == "Tmax"] <- "Tmin"
marbdatetempminna <- rbind(marbdatetempmin, marbwthrna)
plot(marbdatetempminna$Tmin ~ marbdatetempminna$Date, type = "b", xlab = "Date", ylab = "Minimum Daily Temperature (°C)", main = "Marbles Minimum Daily Temperature")

# South Bristols
sbrsdaymet <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/SBrsDaymet11193_lat_34.5937_lon_-115.6823.csv", sep = ",", header = TRUE)
# It rained 2 mm in the South Bristols on 7/25. 
sbrsdaymet$Date <- as.character(sbrsdaymet$Date)
sbrsdaymet$Date <- as.Date(sbrsdaymet$Date)
sbrsdaymetcut <- sbrsdaymet[(sbrsdaymet$Date >= as.Date("2019-06-05") & sbrsdaymet$Date <= as.Date("2019-06-15")) | (sbrsdaymet$Date >= as.Date("2019-06-19") & sbrsdaymet$Date <= as.Date("2019-07-28")) | (sbrsdaymet$Date >= as.Date("2019-08-02") & sbrsdaymet$Date <= as.Date("2019-08-11")), ]
sbrswthrna <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/BrsSNAweatherenvironmentalvariables.csv", sep = ",", header = TRUE)
sbrswthrna$Date <- as.Date(sbrswthrna$Date)
sbrsdatetempmax <- data.frame(sbrsdaymetcut$Date, sbrsdaymetcut$tmax..deg.c.)
names(sbrsdatetempmax)[names(sbrsdatetempmax) == "sbrsdaymetcut.Date"] <- "Date"
names(sbrsdatetempmax)[names(sbrsdatetempmax) == "sbrsdaymetcut.tmax..deg.c."] <- "Tmax"
sbrsdatetempmaxna <- rbind(sbrsdatetempmax, sbrswthrna)
plot(Tmax ~ Date, data = sbrsdatetempmaxna, type = "b", xlab = "Date", ylab = "Maximum Daily Temperature (°C)", main = "South Bristols Maximum Daily Temperature")
# SBrs min temp SHOULD DO


# CMPR
cmprdaymet <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/CMPR11193_lat_35.2795_lon_-115.099.csv", header = TRUE, sep = ",")
# Rained 24 mm on 8/8
cmprdaymet$Date <- as.character(cmprdaymet$Date)
cmprdaymet$Date <- as.Date(cmprdaymet$Date)
cmprdaymetcut <- cmprdaymet[(cmprdaymet$Date >= as.Date("2019-05-24") & cmprdaymet$Date <= as.Date("2019-05-30")) | (cmprdaymet$Date >= as.Date("2019-06-21") & cmprdaymet$Date <= as.Date("2019-06-26")) | (cmprdaymet$Date >= as.Date("2019-07-12") & cmprdaymet$Date <= as.Date("2019-07-14")) | (cmprdaymet$Date >= as.Date("2019-07-31") & cmprdaymet$Date <= as.Date("2019-08-03")) | (cmprdaymet$Date >= as.Date("2019-08-06") & cmprdaymet$Date <= as.Date("2019-08-14")), ]

# Nopah
nopadaymet <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/Nopah11372_lat_36.0105_lon_-116.1015.csv", header = TRUE, sep = ",")
# 7/25 42 mm precipitation
nopadaymet$Date <- as.character(nopadaymet$Date)
nopadaymet$Date <- as.Date(nopadaymet$Date)
nopadaymetcut <- nopadaymet[(nopadaymet$Date >= as.Date("2019-06-07") & nopadaymet$Date <= as.Date("2019-06-12")) | (nopadaymet$Date >= as.Date("2019-06-26") & nopadaymet$Date <= as.Date("2019-07-07")) | (nopadaymet$Date >= as.Date("2019-07-17") & nopadaymet$Date <= as.Date("2019-07-28")) | (nopadaymet$Date >= as.Date("2019-07-30") & nopadaymet$Date <= as.Date("2019-08-06")), ]



```

I then saw if temperature was different between months for the different mountain ranges.

```{r}
# Marbles max temperature
marbdatetempmaxna$Month <- as.numeric(format(marbdatetempmaxna$Date, "%m"))
marbdatetempmaxna$Month <- as.factor(marbdatetempmaxna$Month)
marbmaxtempmonth <- aov(Tmax ~ Month, data = marbdatetempmaxna)
summary.lm(marbmaxtempmonth)
# Multiple R-squared:  0.9489,	Adjusted R-squared:  0.9444 
# F-statistic: 210.4 on 3 and 34 DF,  p-value: < 2.2e-16
TukeyHSD(marbmaxtempmonth)
#           diff        lwr       upr     p adj
# 6-5 14.4142857 12.4761944 16.352377 0.0000000
# 7-5 17.8642857 15.7777156 19.950856 0.0000000
# 8-5 17.1309524 14.7753360 19.486569 0.0000000
# 7-6  3.4500000  1.7214501  5.178550 0.0000307
# 8-6  2.7166667  0.6714189  4.761914 0.0054639
# 8-7 -0.7333333 -2.9197952  1.453129 0.8018217
# 

```

I tested for a causal relationship between the environmental variables and the rate of bighorn visitation by day. First, in excel, I compiled all of the cut daymet data for the different mountain ranges, and also added the NA data for the days that were not covered by cameras for plotting purposes.



```{r}
allrangedaymetna <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/AllRangesDaymetNA.csv", header = TRUE, sep = ",")
allrangedaymetna$NDVI <- NA
allrangedaymetna[1, 13] <- 0.106075
allrangedaymetna[2, 13] <- 0.098809
allrangedaymetna[3, 13] <- 0.110244
allrangedaymetna[4, 13] <- 0.118357
allrangedaymetna[5, 13] <- 0.028228
allrangedaymetna[6, 13] <- 0.026557
allrangedaymetna[7, 13] <- 0.104746
allrangedaymetna[8, 13] <- 0.108609
allrangedaymetna[9, 13] <- 0.115012
allrangedaymetna[10, 13] <- 0.105015
allrangedaymetna[11, 13] <- 0.120782
allrangedaymetna[12, 13] <- 0.102703
allrangedaymetna[13, 13] <- 0.081198
allrangedaymetna[14, 13] <- 0.106395
allrangedaymetna[15, 13] <- 0.109952
allrangedaymetna[16, 13] <- 0.106986
allrangedaymetna[17, 13] <- 0.105554
allrangedaymetna[18, 13] <- 0.112115
allrangedaymetna[19, 13] <- 0.11176
allrangedaymetna[20, 13] <- 0.10329
allrangedaymetna[21, 13] <- 0.108741
allrangedaymetna[22, 13] <- 0.080393
allrangedaymetna[23, 13] <- 0.100699
# MISSING 7/12 [24, 13] APPEARS dataset is missing the metadata in the .tif file for this day
allrangedaymetna[25, 13] <- 0.101607
allrangedaymetna[26, 13] <- 0.100402
allrangedaymetna[27, 13] <- 0.114127
allrangedaymetna[28, 13] <- 0.100427
allrangedaymetna[29, 13] <- 0.106903
allrangedaymetna[30, 13] <- 0.05947
allrangedaymetna[31, 13] <- 0.00414
allrangedaymetna[32, 13] <- 0.076006
allrangedaymetna[33, 13] <- 0.103137
allrangedaymetna[34, 13] <- 0.109912
allrangedaymetna[35, 13] <- 0.100546
allrangedaymetna[36, 13] <- 0.088859
allrangedaymetna[37, 13] <- 0.103131
allrangedaymetna[38, 13] <- 0.110494
```

Another attempt at NDVI

```{r}
# A = NIR = 2
library("LSRS")
library(raster)
marb712b1 <- raster("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/Marb712b1_MOD09GQ.006_sur_refl_b01_1_doy2019193_aid0001.tif")
marb712b2 <- raster("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/Marb712b2_MOD09GQ.006_sur_refl_b02_1_doy2019193_aid0001.tif")
ndvimarb712 <- NDVI(a = marb712b2, b = marb712b1)
cellStats(x = ndvimarb712, stat = 'mean', na.rm = TRUE)


```

