---
title: "Cleaning the GPS Data"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/AllData3920")
```

### Cutting the GPS Data

## Subsetting by Range & Individual Sheep ID

First, I uploaded the GPS data. I checked for duplicates - there were none. Then, I subset the data for only the mountain ranges that I am interested in: South Bristols (BrsS), Marbles (Marb), Nopah (Nopa), Castle-Piute (CMPR), and Newberry-Rodman-Ord (NROM). Then I subset to only the individual bighorn that I am interested in - those with fix rates of 4 hours or less.

A list of bighorn whose collars are every four hours:

Marbles: BHS_1581 F, BHS_1595 F (died on 8/5), BHS_1588 F
South Bristols: BHS_1580 F (before 2019-07-24 04:00, after 1 hour), 
Nopah: BHS_1738 F
CMPR: none
NROM: BHS_1568 F (went to community center), BHS_1565 F, BHS_1567 F, BHS_1564 F, BHS_1677 F

***am not including NROM BHS_1563M b/c travels into East Ord)
*** SBrs BHS_1580 2019-07-24 4:00 switches from 4 hour to 1 hour, holds steady on that until the end-- COULD INCLUDE IN 1 HOUR FIX Rate


A list of bighorn whose collars are every hour or less:

Marbles: BHS_1589 F (very beginning every four hours?), BHS_1584 F, BHS_1582 F, BHS_1577 F, BHS_1683 M, BHS_1682 M, BHS_1684 M (died on 8/17)
South Bristols: BHS_1688 M, BHS_1685 M, BHS_1689 M, BHS_1687 F, BHS_1686 M, BHS_1580F(after 2019-07-24 04:00)
Nopah: BHS_1736 F, BHS_1740 F, BHS_1739 M
CMPR: BHS_1730 F, BHS_1733 F
NROM: none

```{r}
allgpsdata <- read.csv("AllGPSData20190501to20190912.csv", header = TRUE, sep = ",")
# str(allgpsdata)
library(tidyverse)
summary(duplicated(allgpsdata))
rangegpsdata <- allgpsdata[allgpsdata$MountainRange == "BrsS" | allgpsdata$MountainRange == "Marb" | allgpsdata$MountainRange == "Nopa" | allgpsdata$MountainRange == "CMPR" | allgpsdata$MountainRange == "NROM", ]
# View(rangegpsdata)
# str(rangegpsdata)
rangeidgpsdata <- rangegpsdata[rangegpsdata$KI_Known_ID == "BHS_1581" | rangegpsdata$KI_Known_ID == "BHS_1595" | rangegpsdata$KI_Known_ID == "BHS_1588" | rangegpsdata$KI_Known_ID == "BHS_1580" | rangegpsdata$KI_Known_ID == "BHS_1738" | rangegpsdata$KI_Known_ID == "BHS_1568" | rangegpsdata$KI_Known_ID == "BHS_1565" | rangegpsdata$KI_Known_ID == "BHS_1567" | rangegpsdata$KI_Known_ID == "BHS_1564" | rangegpsdata$KI_Known_ID == "BHS_1677" | rangegpsdata$KI_Known_ID == "BHS_1589" | rangegpsdata$KI_Known_ID == "BHS_1584" | rangegpsdata$KI_Known_ID == "BHS_1582" | rangegpsdata$KI_Known_ID == "BHS_1577" | rangegpsdata$KI_Known_ID == "BHS_1683" | rangegpsdata$KI_Known_ID == "BHS_1682" | rangegpsdata$KI_Known_ID == "BHS_1684" | rangegpsdata$KI_Known_ID == "BHS_1688" | rangegpsdata$KI_Known_ID == "BHS_1685" | rangegpsdata$KI_Known_ID == "BHS_1689" | rangegpsdata$KI_Known_ID == "BHS_1687" | rangegpsdata$KI_Known_ID == "BHS_1686" | rangegpsdata$KI_Known_ID == "BHS_1736" | rangegpsdata$KI_Known_ID == "BHS_1740" | rangegpsdata$KI_Known_ID == "BHS_1739" | rangegpsdata$KI_Known_ID == "BHS_1730" | rangegpsdata$KI_Known_ID == "BHS_1733", ]
# View(rangeidgpsdata)
rangeidgpsdata$TimeLocal <- as.character(rangeidgpsdata$TimeLocal)
rangeidgpsdata$DateTimeLocalTest <- rangeidgpsdata$DateTimeLocal
library(lubridate)
rangeidgpsdata$DateTimeLocalTest <- strptime(rangeidgpsdata$DateTimeLocalTest, "%m/%d/%Y %H:%M")
rangeidgpsdata$DateTimeLocalTestct <- as.POSIXct(rangeidgpsdata$DateTimeLocalTest, tz = "America/Los_Angeles", format = "%m/%d/%Y %H:%M")

```
Subsetting the data to when I had camera data for

I next subset the data to the time periods that I had camera data for. Keeping with my visitation analyses, I cut the GPS data to only include days when the whole 24-hour had camera coverage, and camera coverage was complete.

Marbles:
2019-05-18 - 2019-05-24
2019-06-05 - 2019-06-11
2019-06-19 - 2019-06-22
2019-06-26 - 2019-06-29
2019-07-11 - 2019-07-13
2019-07-18 - 2019-07-24
2019-08-01 - 2019-08-03
2019-08-07 - 2019-08-09

South Bristols:
2019-06-05 - 2019-06-15
2019-06-19 - 2019-07-28
2019-08-02 - 2019-08-11

Nopah:
2019-06-07 - 2019-06-12
2019-06-26 - 2019-07-07
2019-07-17 - 2019-07-28
2019-07-30 - 2019-08-06

Castle Piutes:
2019-05-24 - 2019-05-30
2019-06-21 - 2019-06-26
2019-07-12 - 2019-07-14
2019-07-31 - 2019-08-03
2019-08-06 - 2019-08-14

NROM:
2019-06-28 - 2019-07-01
2019-07-10 - 2019-07-12
2019-07-21
2019-08-09 - 2019-08-18
2019-08-21

```{r}
marbidgpsdata <- rangeidgpsdata[rangeidgpsdata$MountainRange == "Marb", ]
marbidgpsdatacut <- marbidgpsdata[(marbidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-05-18 00:00:00") & marbidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-05-24 23:59:59")) | (marbidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-06-05 00:00:00") & marbidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-06-11 23:59:59")) | (marbidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-06-19 00:00:00") & marbidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-06-22 23:59:59")) | (marbidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-06-26 00:00:00") & marbidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-06-29 23:59:59")) | (marbidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-07-11 00:00:00") & marbidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-07-13 23:59:59")) | (marbidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-07-18 00:00:00") & marbidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-07-24 23:59:59")) | (marbidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-08-01 00:00:00") & marbidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-08-03 23:59:59")) | (marbidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-08-07 00:00:00") & marbidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-08-09 23:59:59")), ]

sbrsidgpsdata <- rangeidgpsdata[rangeidgpsdata$MountainRange == "BrsS", ]
sbrsidgpsdatacut <- sbrsidgpsdata[(sbrsidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-06-05 00:00:00") & sbrsidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-06-15 23:59:59")) | (sbrsidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-06-19 00:00:00") & sbrsidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-07-28 23:59:59")) | (sbrsidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-08-02 00:00:00") & sbrsidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-08-11 23:59:59")), ]

cmpridgpsdata <- rangeidgpsdata[rangeidgpsdata$MountainRange == "CMPR", ]
cmpridgpsdatacut <- cmpridgpsdata[(cmpridgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-05-24 00:00:00") & cmpridgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-05-30 23:59:59")) | (cmpridgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-06-21 00:00:00") & cmpridgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-06-26 23:59:59")) | (cmpridgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-07-12 00:00:00") & cmpridgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-07-14 23:59:59")) | (cmpridgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-07-31 00:00:00") & cmpridgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-08-03 23:59:59")) | (cmpridgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-08-06 00:00:00") & cmpridgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-08-14 23:59:59")), ]

nopaidgpsdata <- rangeidgpsdata[rangeidgpsdata$MountainRange == "Nopa", ]
nopaidgpsdatacut <- nopaidgpsdata[(nopaidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-06-07 00:00:00") & nopaidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-06-12 23:59:59")) | (nopaidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-06-26 00:00:00") & nopaidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-07-07 23:59:59")) | (nopaidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-07-17 00:00:00") & nopaidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-07-28 23:59:59")) | (nopaidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-07-30 00:00:00") & nopaidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-08-06 23:59:59")), ]

nromidgpsdata <- rangeidgpsdata[rangeidgpsdata$MountainRange == "NROM", ]
nromidgpsdatacut <- nromidgpsdata[(nromidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-06-28 00:00:00") & nromidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-07-01 23:59:59")) | (nromidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-07-10 00:00:00") & nromidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-07-12 23:59:59")) | (nromidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-07-21 00:00:00") & nromidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-07-21 23:59:59")) | (nromidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-08-09 00:00:00") & nromidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-08-18 23:59:59")) | (nromidgpsdata$DateTimeLocalTestct >= as.POSIXct("2019-08-21 00:00:00") & nromidgpsdata$DateTimeLocalTestct <= as.POSIXct("2019-08-21 23:59:59")), ]

allidgpsdatacut <- rbind(marbidgpsdatacut, sbrsidgpsdatacut, cmpridgpsdatacut, nopaidgpsdatacut, nromidgpsdatacut)
# I removed the entries with missing latitudes or longitudes from the dataset.
allidgpsdatacut1 <- allidgpsdatacut[is.na(allidgpsdatacut$Latitude) == FALSE, ]
allidgpsdatacut2 <- allidgpsdatacut1[is.na(allidgpsdatacut1$Longitude) == FALSE, ]

#Removed GPS points for after 1595 died. RIP.
not1595 <- allidgpsdatacut2[allidgpsdatacut2$KI_Known_ID != "BHS_1595", ]
is1595 <- allidgpsdatacut2[allidgpsdatacut2$KI_Known_ID == "BHS_1595", ]
is1595cut <- is1595[(is1595$DateTimeLocalTestct < as.POSIXct("2019-08-05 00:00:00")), ]
allidgpsdatacut3 <- rbind(not1595, is1595cut)

```
Subsetting to every hour and four hour

To remove the locations that were taken at 20 minute fix rates, I subset by the rows where the minutes were between 55 and 10. The number of rows before subsetting was 21172 while the number of rows after subsetting was 17691, demonstrating that rows with locations taken not close to the hour mark were removed. 


```{r}

# Removing the 20 min fixes
# rangeidhour includes collars on both 1-hour and 4-hour fix rates
rangeidhour <- subset(allidgpsdatacut3, format(allidgpsdatacut3$DateTimeLocalTest, "%M") %in% c('00', '01', '02', '03', '04', '05', '06', '07', '08', '09', '10', '55', '56', '57', '58', '59', '54', '53', '52', '51', '50'))
## I double checked the sheep that are every four hours to make sure that they are on a 0-4-8-12-16-20-24 schedule. They are all.
# rangeidfourhour <- subset(rangeidhour, format(rangeidhour$DateTimeLocalTest, "%H") %in% c('00', '04', '08', '12', '16', '20', '24'))
# Including every hour as ~18,000 gps observations, including only every four hour as ~6,000 observations. Both include 28 individual sheep. 


## Will need to do the amount of time between points for both time schedules before doing any of my 4 gps movement types.


```

### Cleaning the GPS Data

## Mapping the GPS Data

I mapped the GPS data to be able to visually check how cleaning the data changed the sheep movement tracks.

```{r}
# Making an individually identifiable ID for each point
rangeidhour$IndPoint <- paste(rangeidhour$KI_Known_ID, rangeidhour$DateTimeLocalTestct)

library(maptools)
plotattempt <- SpatialPointsDataFrame(rangeidhour[6:5], rangeidhour[21])
plot(plotattempt)

# library(adehabitatLT)
# ltrangeidhour <- as.ltraj(xy = rangeidhour[, c("Latitude", "Longitude")], date = rangeidhour$DateTimeLocalTestct, id = rangeidhour$IndPoint)
# View(ltrangeidhour)

# data.lt_Desert <- as.ltraj(xy = DesertDataDanielleNoDuplicates[, c("UTME", "UTMN")], date = DesertDataDanielleNoDuplicates$DateTimeC, id = DesertDataDanielleNoDuplicates$ID)
#maybe need to convert to UTM first?

library(moveHMM)

names(rangeidhour)[names(rangeidhour) == "KI_Known_ID"] <- "ID"
rangeidhour1 <- rangeidhour[with(rangeidhour, order(ID, DateTimeLocalTestct)), ]
prangeidhour <- prepData(rangeidhour1, type = "LL", coordNames = c("Longitude", "Latitude"))
plot(prangeidhour, compact = T)

library(ggmap)
# plotSat(prangeidhour, zoom = 8)
# Doesn't work because need Google API key????

# Marbles before cleaning
marbidhour <- rangeidhour1[rangeidhour1$MountainRange == "Marb", ]
pmarbidhour <- prepData(marbidhour, type = "LL", coordNames = c("Longitude", "Latitude"))
plot(pmarbidhour, compact = T)
nrow(pmarbidhour)
#5670 points for Marbles before cleaning.

# South Bristols before cleaning
sbrsidhour <- rangeidhour1[rangeidhour1$MountainRange == "BrsS", ]
psbrsidhour <- prepData(sbrsidhour, type = "LL", coordNames = c("Longitude", "Latitude"))
plot(psbrsidhour, compact = T)
nrow(psbrsidhour)
# 7533 points for SBrs before cleaning.

#Nopah before cleaning
nopaidhour <- rangeidhour1[rangeidhour1$MountainRange == "Nopa", ]
pnopaidhour <- prepData(nopaidhour, type = "LL", coordNames = c("Longitude", "Latitude"))
plot(pnopaidhour, compact = T)
nrow(pnopaidhour)
#2655 points for Nopa before cleaning.

#Castle Piute before cleaning
cmpridhour <- rangeidhour1[rangeidhour1$MountainRange == "CMPR", ]
pcmpridhour <- prepData(cmpridhour, type = "LL", coordNames = c("Longitude", "Latitude"))
plot(pcmpridhour, compact = T)
nrow(pcmpridhour)
#1135 points for CMPR before cleaning.

# NROM before cleaning
nromidhour <- rangeidhour1[rangeidhour1$MountainRange == "NROM", ]
pnromidhour <- prepData(nromidhour, type = "LL", coordNames = c("Longitude", "Latitude"))
plot(pnromidhour, compact = T)
nrow(pnromidhour)
# 698 locations for NROM before cleaning.

```
Before looking at HDOP, I first deleted points considered to have location error based on their impossible elevation data (D'Eon et al. 2002). The range of elevations possible in each mountain range was calculated by drawing a polygon around each mountain range outside of CDFW's bighorn population range maps using Caltopo. All points outside of these possible range were eliminated.

Using CalTopo: 
Marbles: 597' - 3840' (181 m - 1171 m)
South Bristols: 703' - 3269' (214 m - 997 m)
NROM: 1800'-6095' (548 m - 1858 m)
Castle Piute: 4000'-5185' (1219 m - 1580 m)
Nopah: 1392'-5249' (424 m - 1599 m)

```{r}
# Marbles
boxplot(pmarbidhour$Altitude)
pmarbidhouralt <- pmarbidhour[(pmarbidhour$Altitude >= 181) & (pmarbidhour$Altitude <= 1171), ]
nrow(pmarbidhouralt)
# Marbles now have 5475 points, instead of 5670 points.
plot(pmarbidhouralt, compact = T)

# South Bristols
boxplot(psbrsidhour$Altitude)
psbrsidhouralt <- psbrsidhour[(psbrsidhour$Altitude >= 214) & (psbrsidhour$Altitude <= 997), ]
nrow(psbrsidhouralt)
# South Bristols now has 7196 points, before had 7533 points.
plot(psbrsidhouralt, compact = T)

#Nopah
boxplot(pnopaidhour$Altitude)
pnopaidhouralt <- pnopaidhour[(pnopaidhour$Altitude >= 424) & (pnopaidhour$Altitude <= 1599), ]
nrow(pnopaidhouralt)
# Nopah now has 2390 points, before had 2655 points.
plot(pnopaidhouralt, compact = T)

#Castle Piutes
boxplot(pcmpridhour$Altitude)
pcmpridhouralt <- pcmpridhour[(pcmpridhour$Altitude >= 1219) & (pcmpridhour$Altitude <= 1580), ]
nrow(pcmpridhouralt)
# Castle Piute now has 1075 points, before had 1135 points.
plot(pcmpridhouralt, compact = T)

# NROM
boxplot(pnromidhour$Altitude)
pnromidhouralt <- pnromidhour[(pnromidhour$Altitude >= 548) & (pnromidhour$Altitude <= 1858), ]
nrow(pnromidhouralt)
# NROM has 672 points, before had 698 points.
plot(pnromidhouralt, compact = T)
```

I then eliminated the points with a DOP above 10 (Lewis et al. 2007), as these points may not be spatially accurate due to minimum distribution of the satellites locating the collar.


```{r}
# Marbles
boxplot(pmarbidhouralt$DOP)
dopmarb <- pmarbidhouralt[pmarbidhouralt$DOP <= 10, ]
nrow(dopmarb)
# Marb now has 5420 points, before had 5475.
plot(dopmarb, compact = T)

# South Bristols
boxplot(psbrsidhouralt$DOP)
dopsbrs <- psbrsidhouralt[psbrsidhouralt$DOP <= 10, ]
nrow(dopsbrs)
# SBrs now has 7150 points, before had 7196.
plot(dopsbrs, compact = T)

# Nopah
boxplot(pnopaidhouralt$DOP)
dopnopa <- pnopaidhouralt[pnopaidhouralt$DOP <= 10, ]
nrow(dopnopa)
# Nopah now has 2371 points, before had 2390.
plot(dopnopa, compact = T)

# Castle Piute
boxplot(pcmpridhouralt$DOP)
dopcmpr <- pcmpridhouralt[pcmpridhouralt$DOP <= 10, ]
nrow(dopcmpr)
# Castle Piute now has 1065 points, before had 1075 points.
plot(dopcmpr, compact = T)

# NROM
boxplot(pnromidhouralt$DOP)
dopnrom <- pnromidhouralt[pnromidhouralt$DOP <= 10, ]
nrow(dopnrom)
# NROM now has 669 points, before had 672.
plot(dopnrom, compact = T)
```

I next examined the relative fix rate by mountain range to check for topographic bias. Since all analyses will be between comparable collars on the same fix schedule, I first separated the fixes back into 1 hour fix rate and 4 hour fix rate. I did subset the 1 hours fix rate collars to increase the 4 hour fix rate sample size. Sheep 1589 switched from 4 hour fix rate to 1 hour fix rate on 2019-06-15 16:00:00.

A lot of the basic, back-of-a-napkin calculations I did on a separate google sheet.


```{r}
dopall <- rbind(dopmarb, dopsbrs, dopcmpr, dopnopa, dopnrom)
# The points after BHS_1595 died are already eliminated from the dataset
# I went back and removed BHS_1692 from the dataset because he went over to East Ord for such a large percentage of the time and I am not confident in my absence of water sources in East Ord.

##Subsetting by sheep to differentiate between sheep with 4-hour collars and sheep with every-hour collars

# SBrs BHS_1580 2019-07-24 4:00 switches from 4 hour to 1 hour, holds steady on that until the end-- COULD INCLUDE IN 1 HOUR FIX Rate

# First differentiated between BHS_1589, BHS_1580 and all other sheep
just1589 <- dopall[dopall$ID == "BHS_1589", ]
dopall1 <- dopall[(dopall$ID != "BHS_1589") & (dopall$ID != "BHS_1580"), ]
just1589before <- just1589[just1589$DateTimeLocalTestct < as.POSIXct("2019-06-15 16:00:00"), ]
just1589after <- just1589[just1589$DateTimeLocalTestct >= as.POSIXct("2019-06-15 16:00:00"), ]
just1580 <- dopall[dopall$ID == "BHS_1580", ]
just1580before <- just1580[just1580$DateTimeLocalTestct < as.POSIXct("2019-07-24 04:00:00"), ]
just1580after <- just1580[just1580$DateTimeLocalTestct >= as.POSIXct("2019-07-24 04:00:00"), ]
# 1 hour fix rate
dopall2 <- dopall1[dopall1$ID == "BHS_1584" |  dopall1$ID == "BHS_1582" | dopall1$ID == "BHS_1577" | dopall1$ID == "BHS_1683" | dopall1$ID == "BHS_1682" | dopall1$ID == "BHS_1684" | dopall1$ID == "BHS_1688" | dopall1$ID == "BHS_1685" | dopall1$ID == "BHS_1689" | dopall1$ID == "BHS_1687" | dopall1$ID == "BHS_1686" | dopall1$ID == "BHS_1736" | dopall1$ID == "BHS_1740" | dopall1$ID == "BHS_1739" | dopall1$ID == "BHS_1730" | dopall1$ID == "BHS_1733", ]
cln1hrfix <- rbind(dopall2, just1589after, just1580after)
# 4 hour fix rate
dopall3 <- dopall1[dopall1$ID == "BHS_1581" | dopall1$ID == "BHS_1595" | dopall1$ID == "BHS_1588" | dopall1$ID == "BHS_1738" | dopall1$ID == "BHS_1568" | dopall1$ID == "BHS_1565" | dopall1$ID == "BHS_1567" | dopall1$ID == "BHS_1564" | dopall1$ID == "BHS_1677", ]
clan4hrfix <- rbind(dopall3, just1589before, just1580before)
# This only includes sheep that are set to 4 hours fix rate. But, can subset 1 hour fix rate to every four hours and include that too.
clan4hrfix1 <- subset(clan4hrfix, format(clan4hrfix$DateTimeLocalTest, "%H") %in% c('00', '04', '08', '12', '16', '20', '24'))
onetofour <- subset(cln1hrfix, format(cln1hrfix$DateTimeLocalTest, "%H") %in% c('00', '04', '08', '12', '16', '20', '24'))
clean4hrfixall <- rbind(clan4hrfix1, onetofour)

## Counting the difference between the actual number of observations and what number there should be if there is assuming a 100% fix rate and no deletion of points while data cleaning

# 1 hr fix rate

# Marbles: 
nrow(cln1hrfix[cln1hrfix$MountainRange == "Marb", ])
# There are 4758 fixes for Marb.
# Based on basic math I did in my google doc, there should be 6048 fixes for Marb.
# 4758/6048 = 78.67% fix rate

# South Bristols:
nrow(cln1hrfix[cln1hrfix$MountainRange == "BrsS", ])
# There are 6921 fixes for South Bristols.
# Based on basic math I did in my google doc, there should be 7676 fixes for SBrs.
# 6691/7320 = 90.16% fix rate

# Nopah
nrow(cln1hrfix[cln1hrfix$MountainRange == "Nopa", ])
# There are 2180 fixes for Nopah.
# There should be 2664 fixes.
# 2180/2664 = 81.83% fix rate

# CMPR
nrow(cln1hrfix[cln1hrfix$MountainRange == "CMPR", ])
# There are 1065 fixes for CMPR.
# There should be 1392 fixes. 
# 1065/1392 = 76.51% fix rate.

# All ranges together
nrow(cln1hrfix)
# Total Actual = 14924 fixes
# Total fixes there should be: 17780
# 14694/17424= 83.94% fix rate



# 4 hour fix rate

# Marbles
nrow(clean4hrfixall[clean4hrfixall$MountainRange == "Marb", ])
# There are 2067 fixes for the Marbles.
# Total fixes there should be: 2262
# 2067/2262 = 91.38% fix rate

# South Bristols
nrow(clean4hrfixall[clean4hrfixall$MountainRange == "BrsS", ])
# There are 1964 fixes for the South Bristols.
# Total fixes there should be: 2196
# 2137/2196 = 89.44%

# Nopah
nrow(clean4hrfixall[clean4hrfixall$MountainRange == "Nopa", ])
# There are 795 for the Nopahs.
# Total fixes there should be: 888
# 795/888 = 89.53% fix rate

# CMPR
nrow(clean4hrfixall[clean4hrfixall$MountainRange == "CMPR", ])
# 325 points for CMPR.
# Total fixes there should be: 348
# 93.34% fix rate

# NROM
nrow(clean4hrfixall[clean4hrfixall$MountainRange == "NROM", ])
# 533 points for NROM.
# Total fixes there should be: 570
# 93.51% fix rate

# All mtn ranges together except NROM:
nrow(clean4hrfixall[clean4hrfixall$MountainRange == "Marb" | clean4hrfixall$MountainRange == "BrsS" | clean4hrfixall$MountainRange == "Nopa" | clean4hrfixall$MountainRange == "CMPR",])
# 5151 points for Marb, BrsS, Nopa, CMPR.
# Total fixes there should be: 5694
# 5151/5694 = 90.46% fix rate

# All mtn ranges together including NROM:
nrow(clean4hrfixall)
# 5684 points total
# Total fixes there should be: 6264
# 5684/6264 = 90.74% fix rate
```
I then started to calculate my four measures of GPS movement: step length, turning angle, distance to water, angle of movement relative to water. I decided to start with 1 hr, then do 4 hr. To start with, I calculated turning angle.

- if turning angle is different between different sheep- DONE
- if turning angle is different by month- DONE
- if turning angle is different by mountain range - DONE
- if turning angle is different by time-of-day

```{r}
library(moveHMM)
# 1 hr fix
names(cln1hrfix)[names(cln1hrfix) == "y"] <- "Latitude"
names(cln1hrfix)[names(cln1hrfix) == "x"] <- "Longitude"
cln1hrfix1 <- cln1hrfix[with(cln1hrfix, order(ID, DateTimeLocalTestct)), ]
pcln1hrfix <- prepData(cln1hrfix1, type = "LL", coordNames = c("Longitude", "Latitude"))
# Upon examination, these turning angles are in radians. The min and max are pi, which makes sense because it is impossible to have a turning angle greater than 180 degrees since the angle that the animal turned should be assumed to be the smaller one (why would I turn 270 degrees when I can only turn 90?). Since for this metric, all I am looking at is the tortuosity of the track, I think I should make every radian angle positive. After all, I do not care about whether the sheep turned left or right.
pcln1hrfix$positangle <- abs(pcln1hrfix$angle)
library(amt)
pcln1hrfix$positangledegree <- as_degree(pcln1hrfix$positangle)
# Angles between sheep are already deleted because of how prepData function worked
# Need to make sure that points before and after a given point were actually taken an hour before and an hour after.
pcln1hrfix$lagtime <- (pcln1hrfix$DateTimeLocalTestct - lag(pcln1hrfix$DateTimeLocalTestct))
pcln1hrfix$leadtime <- (lead(pcln1hrfix$DateTimeLocalTestct) - pcln1hrfix$DateTimeLocalTestct)
pcln1hrfix$lagtime <- as.numeric(pcln1hrfix$lagtime)
pcln1hrfix$leadtime <- as.numeric(pcln1hrfix$leadtime)
pcln1hrfix1 <- pcln1hrfix[((pcln1hrfix$lagtime >= 50 & pcln1hrfix$lagtime <= 70) & (pcln1hrfix$leadtime >= 50 & pcln1hrfix$leadtime <= 70)), ]



hist(pcln1hrfix1$positangledegree, xlab = "Angle Degree", main = "Frequency of Turning Angles at 1 Hr Fix Rate for All Bighorn (n = 12964)")
pcln1hrfix1$Month <- month(pcln1hrfix1$DateTimeLocalTestct)
pcln1hrfix1$MountainRange <- droplevels(pcln1hrfix1$MountainRange)
# If turning angle is different by mountain range
boxplot(positangledegree ~ MountainRange, data = pcln1hrfix1, ylab = "Turning Angle", main = "Turning Angle by Mountain Range")
turnangleMtnRange <- aov(positangledegree ~ MountainRange, data = pcln1hrfix1)
summary.lm(turnangleMtnRange)
# Multiple R-squared:  0.0003166, Adjusted R-squared:  8.509e-05
# F-statistic: 1.368 on 3 and 12955 DF,  p-value: 0.2506
TukeyHSD(turnangleMtnRange)
#                 diff       lwr      upr     p adj
# CMPR-BrsS  1.7289660 -3.340182 6.798114 0.8171345
# Marb-BrsS -1.7841234 -4.683856 1.115610 0.3895996
# Nopa-BrsS -0.6429978 -4.457247 3.171252 0.9727806
# Marb-CMPR -3.5130893 -8.759514 1.733336 0.3129495
# Nopa-CMPR -2.3719637 -8.174120 3.430193 0.7196964
# Nopa-Marb  1.1411256 -2.905751 5.188002 0.8873445
# Turning angle is not significantly different between any mountain ranges.

# If turning angle is significantly different by individual sheep
boxplot(positangledegree ~ ID, data = pcln1hrfix1, ylab = "Turning Angle", main = "Turning Angle by Individual Bighorn")
turnangleid <- aov(positangledegree ~ ID, data = pcln1hrfix1)
summary.lm(turnangleid)
# Multiple R-squared:  0.00476, Adjusted R-squared:  0.003453
# F-statistic: 3.641 on 17 and 12941 DF,  p-value: 5.313e-07
TukeyHSD(turnangleid)
# Only bighorn combos that were significantly different were: 1686-1584, 1686-1682, 1687-1686, 1688-1686, 1740-1686, 1740-1739.

# If turning angle is different by month
# Because there was no significant difference between mountain ranges, I did not divide up these results by mountain range.
boxplot(positangledegree ~ Month, data = pcln1hrfix1)
pcln1hrfix1$Month <- as.factor(pcln1hrfix1$Month)
turnanglemonth <- aov(positangledegree ~ Month, data = pcln1hrfix1)
summary.lm(turnanglemonth)
# Multiple R-squared:  0.0009122, Adjusted R-squared:  0.0006808
# F-statistic: 3.943 on 3 and 12955 DF,  p-value: 0.008015
TukeyHSD(turnanglemonth)
#           diff        lwr        upr     p adj
# 6-5 -2.3332716  -9.200252  4.5337086 0.8188386
# 7-5 -5.5160170 -12.346045  1.3140114 0.1613233
# 8-5 -5.2894958 -12.395589  1.8165977 0.2227202
# 7-6 -3.1827453  -6.071774 -0.2937171 0.0240622
# 8-6 -2.9562242  -6.448181  0.5357326 0.1302038
# 8-7  0.2265211  -3.192197  3.6452390 0.9982468
# There was only a significant different in turning angle between June and July. The fact that there wasn't a significant between farther apart months brings into question the relevance of this significance.

```

I next calculated the shortest topographic distance between points. This is basically step length, accounting for elevation.

```{r}

#Nopah

# Setting up the elevation raster
library(raster)
library(sp)
library(rgdal)
# nopahelevation <- raster("C:/Users/Danielle Glass/Downloads/nopahelevation.tif")
# proj4string(nopahelevation) <- CRS("+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")
# plot(nopahelevation, compact = TRUE)
library(rgdal)
# tnopahelevation <- projectRaster(from = nopahelevation, crs = ("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))
# plot(tnopahelevation)
# nopahelevationattempt2 <- raster("C:/Users/Danielle Glass/Downloads/NopahElevation1.tif")
# proj4string(nopahelevationattempt2) <- CRS("+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")

nopahelevationattempt3 <- raster("C:/Users/dmg85/Downloads/Sheepies3-master/output_srtm.tif")
#
# Attempt below didn't work, but one above did?
# nopahelevationproj <- projectRaster(nopahelevation, crs = "+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")

# WGS 84

# Setting up the SpatialPointsObjects
onehrnopah <- pcln1hrfix[pcln1hrfix$MountainRange == "Nopa", ]
onehrnopah$ID <- droplevels(onehrnopah$ID)
coordsnopa <- as.data.frame(onehrnopah[ ,c("x", "y")])
restodatanopa <- as.data.frame(onehrnopah[ , c(1:3, 6:28)])
crsnopa <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
# crsnopa <- CRS("+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")
nopaspdf <- SpatialPointsDataFrame(coords = coordsnopa, data = restodatanopa, proj4string = crsnopa)
# tnopaspdf <- spTransform(nopaspdf , CRS("+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs"))
# jpeg("tnopaspdf.jpg", width = 8, height = 12, units = "in", res = 180)
# plot(tnopaspdf)
# dev.off()
# View(nopaspdf@data)

library(topoDistance)
topodistnopa <- topoDist(DEM = nopahelevationattempt3, pts = nopaspdf, directions = 8, paths = FALSE, zweight = 1)
View(topodistnopa)
topodistnopa1 <- as.matrix(topodistnopa)
topodistnopa1sub <- topodistnopa1[row(topodistnopa1) == (col(topodistnopa1) - 1)]
View(topodistnopa1sub)
natopodistnopa1sub <- c(NA, topodistnopa1sub)

onehrnopah2 <- cbind(onehrnopah, natopodistnopa1sub)
# Then deleted the wrong distances that were calculated between sheep.
onehrnopah2 <- cbind(rownames(onehrnopah2), data.frame(onehrnopah2, row.names = NULL))
onehrnopah2[685, 30] <- NA
onehrnopah2[1389, 30] <- NA

# Castle Piute

castlepiuteelevation <- raster("C:/Users/dmg85/Downloads/Sheepies3-master/castlepiuteelevation.tif")
onehrcmpr <- pcln1hrfix[pcln1hrfix$MountainRange == "CMPR",]
onehrcmpr$ID <- droplevels(onehrcmpr$ID)
coordscmpr <- as.data.frame(onehrcmpr[ ,c("x", "y")])
restodatacmpr <- as.data.frame(onehrcmpr[ , c(1:3, 6:28)])
crscmpr <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
cmprspdf <- SpatialPointsDataFrame(coords = coordscmpr, data = restodatacmpr, proj4string = crscmpr)
topodistcmpr <- topoDist(DEM = castlepiuteelevation, pts = cmprspdf, directions = 8, paths = FALSE, zweight = 1)
topodistcmpr1 <- as.matrix(topodistcmpr)
topodistcmpr1sub <- topodistcmpr1[row(topodistcmpr1) == (col(topodistcmpr1) - 1)]
# View(topodistcmpr1sub)
natopodistcmpr1sub <- c(NA, topodistcmpr1sub)
onehrcmpr2 <- cbind(onehrcmpr, natopodistcmpr1sub)
onehrcmpr2 <- cbind(rownames(onehrcmpr2), data.frame(onehrcmpr2, row.names = NULL))
onehrcmpr2[546, 30] <- NA

# Marbles
marbleselevation <- raster("C:/Users/dmg85/Downloads/Sheepies3-master/marbleselevation.tif")
onehrmarb <- pcln1hrfix[pcln1hrfix$MountainRange == "Marb", ]
onehrmarb$ID <- droplevels(onehrmarb$ID)
coordsmarb <- as.data.frame(onehrmarb[ , c("x", "y")])
restodatamarb <- as.data.frame(onehrmarb[ , c(1:3, 6:28)])
crsmarb <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
marbspdf <- SpatialPointsDataFrame(coords = coordsmarb, data = restodatamarb, proj4string = crsmarb)
topodistmarb <- topoDist(DEM = marbleselevation, pts = marbspdf, directions = 8, paths = FALSE, zweight = 1)
topodistmarb1 <- as.matrix(topodistmarb)
topodistmarb1sub <- topodistmarb1[row(topodistmarb1) == (col(topodistmarb1) - 1)]
# View(topodistmarb1sub)
natopodistmarb1sub <- c(NA, topodistmarb1sub)
onehrmarb2 <- cbind(onehrmarb, natopodistmarb1sub)
onehrmarb2 <- cbind(rownames(onehrmarb2), data.frame(onehrmarb2, row.names = NULL))
onehrmarb2[511, 30] <- NA
onehrmarb2[1108, 30] <- NA
onehrmarb2[1707, 30] <- NA
onehrmarb2[2201, 30] <- NA
onehrmarb2[3086, 30] <- NA
onehrmarb2[3949, 30] <- NA

# South Bristols
sbrselevation <- raster("C:/Users/dmg85/Downloads/Sheepies3-master/sbristolselevation.tif")
onehrsbrs <- pcln1hrfix[pcln1hrfix$MountainRange == "BrsS", ]
onehrsbrs$ID <- droplevels(onehrsbrs$ID)
coordsbrs <- as.data.frame(onehrsbrs[ , c("x", "y")])
restodatasbrs <- as.data.frame(onehrsbrs[ , c(1:3, 6:28)])
crsbrs <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
sbrspdf <- SpatialPointsDataFrame(coords = coordsbrs, data = restodatasbrs, proj4string = crsbrs)
topodistsbrs <- topoDist(DEM = sbrselevation, pts = sbrspdf, directions = 8, paths = FALSE, zweight = 1)
topodistsbrs1 <- as.matrix(topodistsbrs)
topodistsbrs1sub <- topodistsbrs1[row(topodistsbrs1) == (col(topodistsbrs1) - 1)]
# View(topodistsbrs1sub)
natopodistsbrs1sub <- c(NA, topodistsbrs1sub)
onehrsbrs2 <- cbind(onehrsbrs, natopodistsbrs1sub)
onehrsbrs2 <- cbind(rownames(onehrsbrs2), data.frame(onehrsbrs2, row.names = NULL))
onehrsbrs2[231, 30] <- NA
onehrsbrs2[1579, 30] <- NA
onehrsbrs2[2990, 30] <- NA
onehrsbrs2[4322, 30] <- NA
onehrsbrs2[5692, 30] <- NA

# Need the row names to be the same in order to rbind to work
names(onehrnopah2)[names(onehrnopah2) == "rownames(onehrnopah2)"] <- "rownumlrgdatset"
names(onehrcmpr2)[names(onehrcmpr2) == "rownames(onehrcmpr2)"] <- "rownumlrgdatset"
names(onehrmarb2)[names(onehrmarb2) == "rownames(onehrmarb2)"] <- "rownumlrgdatset"
names(onehrsbrs2)[names(onehrsbrs2) == "rownames(onehrsbrs2)"] <- "rownumlrgdatset"

names(onehrnopah2)[names(onehrnopah2) == "natopodistnopa1sub"] <- "stplengthtopo"
names(onehrcmpr2)[names(onehrcmpr2) == "natopodistcmpr1sub"] <- "stplengthtopo"
names(onehrmarb2)[names(onehrmarb2) == "natopodistmarb1sub"] <- "stplengthtopo"
names(onehrsbrs2)[names(onehrsbrs2) == "natopodistsbrs1sub"] <- "stplengthtopo"
steplengthallmtn <- rbind(onehrnopah2, onehrcmpr2, onehrmarb2, onehrsbrs2)


# library(dplyr)
# BrsSlongtimedif <- BrsScutobslong %>%
#   arrange(KI_Known_ID, DateTimect) %>%
#   group_by(KI_Known_ID) %>%
#   mutate(Timediff = DateTimect - lag(DateTimect), diff_hours = as.numeric(Timediff, units = 'hours'))

# library(adehabitatLT)
# data.lt.nopah <- as.ltraj(xy = onehrnopah[, c("x", "y")], date = onehrnopah$DateTimeLocalTestct, id = onehrnopah$ID)
# View(data.lt.nopah[[1]])

# data.lt_Desert <- as.ltraj(xy = DesertDataDanielleNoDuplicates[, c("UTME", "UTMN")], date = DesertDataDanielleNoDuplicates$DateTimeC, id = DesertDataDanielleNoDuplicates$ID)

# data.lt<-as.ltraj(xy=data[,c("x","y")], date=data$date, id=data$id)





```

I conducted the preliminary analyses related to step length. I read in the data at the beginning because I had to conduct the previous chunk on a computer more powerful than my laptop. All topographic step lengths are in meters.

- step length different by mountain range
- different by individual sheep
- different by month

```{r}
steplengthallmtn <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory3/Sheepies5/steplengthallmtn3820.csv")
# Need to make sure that points before and after a given point were actually taken an hour before and an hour after.
library(lubridate)
steplengthallmtn$DateTimeLocalTestct <- as.character(steplengthallmtn$DateTimeLocalTestct)
steplengthallmtn$DateTimeLocalTestct <- strptime(steplengthallmtn$DateTimeLocalTestct, "%m/%d/%Y %H:%M")
steplengthallmtn$DateTimeLocalTestct <- as.POSIXct(steplengthallmtn$DateTimeLocalTestct, tz = "America/Los_Angeles", format = "%m/%d/%Y% %H:%M")
steplengthallmtn$lagtime <- (steplengthallmtn$DateTimeLocalTestct - lag(steplengthallmtn$DateTimeLocalTestct))

steplengthallmtn$lagtime <- difftime((lag(steplengthallmtn$DateTimeLocalTestct)), steplengthallmtn$DateTimeLocalTestct, units = "mins")
 
  #(steplengthallmtn$DateTimeLocalTestct - lag(steplengthallmtn$DateTimeLocalTestct))




pcln1hrfix$lagtime <- (pcln1hrfix$DateTimeLocalTestct - lag(pcln1hrfix$DateTimeLocalTestct))
pcln1hrfix$leadtime <- (lead(pcln1hrfix$DateTimeLocalTestct) - pcln1hrfix$DateTimeLocalTestct)
pcln1hrfix$lagtime <- as.numeric(pcln1hrfix$lagtime)
pcln1hrfix$leadtime <- as.numeric(pcln1hrfix$leadtime)
pcln1hrfix1 <- pcln1hrfix[((pcln1hrfix$lagtime >= 50 & pcln1hrfix$lagtime <= 70) & (pcln1hrfix$leadtime >= 50 & pcln1hrfix$leadtime <= 70)), ]






# Step length different by mountain range
# The following hist does not include all step lengths, there are a couple in the thousands, max = ~8000 meters.
hist(steplengthallmtn$stplengthtopo, breaks = 80, xlim = c(0, 1500), xlab = "Step Length (m)", main = "Frequency of Step Length Given All Bighorn")
boxplot(stplengthtopo ~ MountainRange, data = steplengthallmtn)


# hist(pcln1hrfix$positangledegree, xlab = "Angle Degree", main = "Frequency of Turning Angles at 1 Hr Fix Rate for All Bighorn (n = 14924)")
# pcln1hrfix$Month <- month(pcln1hrfix$DateTimeLocalTestct)
# pcln1hrfix$MountainRange <- droplevels(pcln1hrfix$MountainRange)
# If turning angle is different by mountain range
# boxplot(positangledegree ~ MountainRange, data = pcln1hrfix, ylab = "Turning Angle", main = "Turning Angle by Mountain Range")
# turnangleMtnRange <- aov(positangledegree ~ MountainRange, data = pcln1hrfix)
# summary.lm(turnangleMtnRange)
# TukeyHSD()
```
This code is what I used before I went into the lab and did it on a computer that could handle the topoDistance function.
```{r}
# Setting up the elevation raster
library(raster)
library(sp)
library(rgdal)
nopahelevation <- raster("C:/Users/Danielle Glass/Downloads/nopahelevation.tif")
proj4string(nopahelevation) <- CRS("+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")
plot(nopahelevation, compact = TRUE)
library(rgdal)
# tnopahelevation <- projectRaster(from = nopahelevation, crs = ("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))
# plot(tnopahelevation)
nopahelevationattempt2 <- raster("C:/Users/Danielle Glass/Downloads/NopahElevation1.tif")
proj4string(nopahelevationattempt2) <- CRS("+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")

nopahelevationattempt3 <- raster("C:/Users/dmg85/Downloads/Sheepies3-master/output_srtm.tif")
# 
# Attempt below didn't work, but one above did?
# nopahelevationproj <- projectRaster(nopahelevation, crs = "+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")

# WGS 84 

# Setting up the SpatialPointsObjects

# This step length calculated automatically from the prepData function of moveHMM does not account for geodetic (the earth being a sphere), or topopgraphy.
pcln1hrfix$stepnotopo <- (pcln1hrfix$step * 1000)

# This one does:
onehrnopah <- pcln1hrfix[pcln1hrfix$MountainRange == "Nopa", ]
onehrnopah$ID <- droplevels(onehrnopah$ID)
coordsnopa <- as.data.frame(onehrnopah[ ,c("x", "y")])
restodatanopa <- as.data.frame(onehrnopah[ , c(1:3, 6:27)])
crsnopa <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
# crsnopa <- CRS("+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs")
nopaspdf <- SpatialPointsDataFrame(coords = coordsnopa, data = restodatanopa, proj4string = crsnopa)
# tnopaspdf <- spTransform(nopaspdf , CRS("+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs"))
jpeg("tnopaspdf.jpg", width = 8, height = 12, units = "in", res = 180)
plot(tnopaspdf)
dev.off()
View(nopaspdf@data)

library(topoDistance)
topodistnopa <- topoDist(DEM = nopahelevationattempt2, pts = tnopaspdf, directions = 4, paths = FALSE, zweight = 1)
View(topodistnopa)
topodistnopa1 <- as.matrix(topodistnopa)
topodistnopa1sub <- topodistnopa1[row(topodistnopa1) == (col(topodistnopa1) - 1)]
View(topodistnopa1sub)
natopodistnopa1sub <- c(NA, topodistnopa1sub)

onehrnopah2 <- cbind(onehrnopah, natopodistnopa1sub)

write.csv(onehrnopah2, "C:/Users/Danielle Glass/Documents/RWorkingDirectory2/Sheepies3/pcln1hrfix3820.csv")
  #onehrnopah$topodistnopahstplength <- topodistnopa1sub


#View(onehrnopah2)


# library(dplyr)
# BrsSlongtimedif <- BrsScutobslong %>%
#   arrange(KI_Known_ID, DateTimect) %>%
#   group_by(KI_Known_ID) %>%
#   mutate(Timediff = DateTimect - lag(DateTimect), diff_hours = as.numeric(Timediff, units = 'hours'))

library(adehabitatLT)
data.lt.nopah <- as.ltraj(xy = onehrnopah[, c("x", "y")], date = onehrnopah$DateTimeLocalTestct, id = onehrnopah$ID)
View(data.lt.nopah[[1]])

# data.lt_Desert <- as.ltraj(xy = DesertDataDanielleNoDuplicates[, c("UTME", "UTMN")], date = DesertDataDanielleNoDuplicates$DateTimeC, id = DesertDataDanielleNoDuplicates$ID)

# data.lt<-as.ltraj(xy=data[,c("x","y")], date=data$date, id=data$id)





```
Reading in the step length incorporating topography csv. NEED TO WORK ON THIS.


```{r}
steplengthallmtn <- read.csv("C:/Users/Danielle Glass/Documents/RWorkingDirectory2/Sheepies3/steplengthallmtn3820.csv")




```

```{r}
library(rosm)
library(prettymapr)

marbles <- makebbox(34.743890,-115.475526,34.521739,-115.699376)

# n, e, s, w
bmaps.plot(marbles, type = "Aerial", zoom = 10)


# sp package
# talk to Lorena

library(moveHMM)


# JUNK CODE FROM ATTEMPTS:
# ggmap(get_stamenmap(bbox = c(left = -115.656813, bottom = 34.648079, right = -115.471502, top = 34.639717), maptype = "terrain", crop = FALSE, zoom = 12))

# install.packages("devtools")
# install.packages("ggmap")
# library(ggmap)
# library(ggplot2)
# library("devtools")
# marblesbasemap <- get_map(location = c(lon = 115.582657, lat = 34.641645), zoom = 11, maptype = 'hybrid', source = 'stamen') 

# bbox <- c(left = 34.639717, bottom = 115.471502, right = 34.648079, top = 115.656813)
# ggmap(get_map(bbox, zoom = 13))

# bbox <- get_stamenmap(bbox = c(left = 34.639717, bottom = 115.471502, right = 34.648079, top = 115.656813), maptype = "terrain", crop = FALSE, zoom = 10)

# Attempt #2
# get_map(location = c(lon = -115.571137, lat = 34.636558), zoom = 10, scale = "auto", maptype = "terrain", source = "google")

```



```{r}
onehrnopah


# library(dplyr)
# BrsSlongtimedif <- BrsScutobslong %>%
#   arrange(KI_Known_ID, DateTimect) %>%
#   group_by(KI_Known_ID) %>%
#   mutate(Timediff = DateTimect - lag(DateTimect), diff_hours = as.numeric(Timediff, units = 'hours'))

library(adehabitatLT)
library(dplyr)
data.lt.nopah <- as.ltraj(xy = onehrnopah[, c("x", "y")], date = onehrnopah$DateTimeLocalTestct, id = onehrnopah$ID)
View(data.lt.nopah[[1]])
for(i in 1:3){data.lt.nopah[[i]]$xend <- lag(data.lt.nopah[[i]]$x)}
for(i in 1:3){data.lt.nopah[[i]]$yend <- lag(data.lt.nopah[[i]]$y)}
nopah1736 <- as.data.frame(data.lt.nopah[[1]])
write.csv(nopah1736, "C:/Users/Danielle Glass/Documents/RWorkingDirectory2/Sheepies3/nopah1736.csv")

# hi
```

